---
title: "Mental Health Services Utilization Among Immigrants in Canada"
author: "ALEXANDER ANYAGRE"
date: "`r Sys.Date()`"
output: html_document
---

```{r init_setup, include=FALSE}
# Initial setup chunk - renamed to avoid duplication
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r warning=FALSE,error=FALSE,message=FALSE}

#Load Required Packages

library(tidyverse)
library(haven)
library(gtsummary)
library(gt)
library(performance)
library(car)
library(nnet)
library(pscl)
library(lavaan)
library(patchwork)
#Data Processing Pipeline

# Load and prepare data
data <- read_dta("D:/Assaya/Assaya/Data/JULY232025.dta")
colnames(data) <- tolower(names(data))
```


```{r warning=FALSE,error=FALSE,message=FALSE}
# Initial data conversion
data_processed <- data %>%
  rename(
    pop_centre_rural = "geodvpsz",
    marital_status = "dhhgms",
    gender = "gender",
    age_group = "dhhgage",
    immigration_status = "sdcfimm",
    years_in_canada = "sdcgres",
    education = "edu_05",
    perceived_mh_compared_prev_year = "gen_08c",
    perceived_life_stress = "gen_07",
    perceived_work_stress = "gen_09",
    self_perceived_mh = "gen_08b",
    depression_interference_12m = "depdint",
    suicide_thought_12m_flag = "depfsyt",
    depression_hospitalized_life = "dep_87",
    depression_consulted_prof_life = "dep_72",
    suicide_ideation_12m = "sui_03",
    suicide_ideation_worst_episode = "sui_01",
    suicide_ideation_life = "sui_02",
    consulted_ment_hlth_12m = "sr1_004j"
  ) %>%
  mutate(
    across(
      starts_with("sr1_"), 
      ~factor(
        case_when(
          .x == 1 ~ "Yes",
          .x == 2 ~ "No",
          TRUE ~ NA_character_
        ),
        levels = c("No", "Yes")
      )
    )
  ) %>%
  mutate(
    pop_centre_rural = factor(
      pop_centre_rural,
      levels = c(1, 2, 3, 4),
      labels = c("Large urban", "Medium urban", "Small urban", "Rural")
    ),
    marital_status = factor(
      marital_status,
      levels = c(1, 2, 3, 4, 5),
      labels = c("Married", "Common-law", "Widowed/Divorced/Separated", "Single", "Other")
    ),
    gender = factor(gender, levels = c(1, 2), labels = c("Male", "Female")),
    age_group = factor(
      age_group,
      levels = c(1, 2, 3, 4, 5, 6, 7, 8),
      labels = c("12-14", "15-17", "18-19", "20-24", "25-29", "30-34", "35-39", "40+")
    ),
    immigration_status = factor(
      immigration_status,
      levels = c(1, 2),
      labels = c("Immigrant", "Non-immigrant")
    ),
    years_in_canada = factor(
      years_in_canada,
      levels = c(1, 2, 3),
      labels = c("0-5 years", "6-10 years", "11+ years")
    ),
    education = factor(
      education,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "Less than high school", "High school", "Trade certificate",
        "College diploma", "University certificate below bachelor",
        "Bachelor's degree", "University above bachelor"
      )
    )
  )
```

#Service Use Flag Creation
```{r warning=FALSE,error=FALSE,message=FALSE, results='asis'}

service_groups <- list(
  professional = c("sr1_004a", "sr1_004b", "sr1_004c", "sr1_004d", "sr1_004e"),
  informal = c("sr1_004f", "sr1_004g", "sr1_004h", "sr1_113", "sr1_116"),
  text = c("sr1_10ad", "sr1_20ad", "sr1_30ad", "sr1_50ad", "sr1_60ad", "sr1_70ad", "sr1_80ad", "sr1100ad"),
  phone = c("sr1_10ab", "sr1_20ab", "sr1_30ab", "sr1_40ab", "sr1_50ab", "sr1_60ab", "sr1_70ab", "sr1_80ab", "sr1100ab", "sr1_116"),
  video = c("sr1_10ac", "sr1_20ac", "sr1_30ac", "sr1_40ac", "sr1_50ac", "sr1_60ac", "sr1_70ac", "sr1_80ac", "sr1100ac"),
  inperson = c("sr1_10aa", "sr1_20aa", "sr1_30aa", "sr1_40aa", "sr1_50aa", "sr1_60aa", "sr1_70aa", "sr1_80aa", "sr1100aa")
)

create_service_flag <- function(data, vars) {
  existing_vars <- vars[vars %in% names(data)]
  if (length(existing_vars) == 0) return(rep(NA, nrow(data)))
  flag <- apply(data[, existing_vars], 1, function(x) any(x == "Yes", na.rm = TRUE))
  factor(ifelse(flag, "Yes", "No"), levels = c("No", "Yes"))
}

data_processed <- data_processed %>%
  mutate(
    professional_service_any = create_service_flag(., service_groups$professional),
    informal_support_any = create_service_flag(., service_groups$informal),
    any_text_sms_service_use = create_service_flag(., service_groups$text),
    any_phone_service_use = create_service_flag(., service_groups$phone),
    any_video_service_use = create_service_flag(., service_groups$video),
    any_inperson_service_use = create_service_flag(., service_groups$inperson)
  )
```

#Analysis Functions
```{r results='asis'}

# Enhanced Service Utilization by Demographics
# Enhanced Service Utilization by Demographics - FIXED VERSION
demo_service_analysis <- function(data) {
  data <- data %>%
    mutate(service_type = case_when(
      professional_service_any == "Yes" & informal_support_any == "Yes" ~ "Both",
      professional_service_any == "Yes" ~ "Professional Only",
      informal_support_any == "Yes" ~ "Informal Only",
      TRUE ~ "None"
    ) %>% factor(levels = c("None", "Professional Only", "Informal Only", "Both")))
  
  create_demo_table <- function(var) {
    # First create summary table by service type
    tbl_by_service <- data %>%
      filter(immigration_status == "Immigrant") %>%
      mutate({{var}} := fct_drop({{var}})) %>%
      tbl_summary(
        by = service_type,
        include = {{var}},
        missing = "no"
      ) %>%
      add_p() %>%
      modify_header(update = list(label ~ "**Variable**"))
    
    # Then create overall summary
    tbl_overall <- data %>%
      filter(immigration_status == "Immigrant") %>%
      mutate({{var}} := fct_drop({{var}})) %>%
      tbl_summary(
        include = {{var}},
        missing = "no"
      )
    
    # Combine tables with proper caption formatting
    tbl_merge(list(tbl_overall, tbl_by_service)) %>%
      modify_caption(paste("**Service Utilization Patterns by", as_label(enquo(var)), "**")) %>%
      modify_footnote(update = everything() ~ NA)
  }
  
  list(
    gender = create_demo_table(gender),
    age = create_demo_table(age_group),
    years_canada = create_demo_table(years_in_canada),
    education = create_demo_table(education)
  )
}
# Professional-specific Barrier Analysis
barrier_analysis <- function(data) {
  barrier_data <- data %>%
    filter(immigration_status == "Immigrant") %>%
    mutate(
      across(c(starts_with("sr1_0")), ~ifelse(.x == "Yes", 1, 0)),
      any_professional = ifelse(professional_service_any == "Yes", 1, 0)
    )
  
  prof_types <- c("Psychiatrist" = "004a", "Family Doctor" = "004b",
                 "Psychologist" = "004c", "Nurse" = "004d", "Social Worker" = "004e")
  
  map(prof_types, ~{
    service_var <- paste0("sr1_", .x)
    barrier_vars <- names(data)[str_detect(names(data), paste0("sr1_", substr(.x, 1, 2), "\\d[a-z]$"))]
    
    frm <- reformulate(barrier_vars, response = service_var)
    glm(frm, data = barrier_data, family = binomial) %>%
      tbl_regression(exponentiate = TRUE) %>%
      modify_header(label ~ "**Barrier**") %>%
      modify_caption(paste("**Barriers Predicting", names(prof_types)[prof_types == .x], "Service Use**"))
  })
}

```

# Longitudinal Mental Health Analysis
```{r warning=FALSE,error=FALSE,message=FALSE}
mh_trajectory_analysis <- function(data) {
  # Safely convert variables and handle missing data
  data <- data %>%
    mutate(
      perceived_mh_compared_prev_year = labelled::to_factor(perceived_mh_compared_prev_year),
      mh_trajectory = case_when(
        is.na(perceived_mh_compared_prev_year) ~ NA_character_,
        perceived_mh_compared_prev_year %in% c("Much better", "Somewhat better") ~ "Improved",
        perceived_mh_compared_prev_year == "About the same" ~ "Stable",
        perceived_mh_compared_prev_year %in% c("Somewhat worse", "Much worse") ~ "Declined",
        TRUE ~ NA_character_
      ) %>% 
        factor(levels = c("Declined", "Stable", "Improved"))
    )
  
  # Prepare modeling data with complete cases only
  model_data <- data %>%
    select(mh_trajectory, professional_service_any, informal_support_any,
           gender, age_group, years_in_canada) %>%
    mutate(across(where(is.factor), fct_drop)) %>%
    drop_na()
  
  # Check for sufficient variation in predictors
  valid_predictors <- c("professional_service_any", "informal_support_any",
                       "gender", "age_group", "years_in_canada") %>%
    keep(~length(unique(model_data[[.x]])) > 1)
  
  # Build results list
  results <- list(
    desc_table = model_data %>%
      tbl_summary(
        include = c(mh_trajectory, professional_service_any, informal_support_any),
        missing = "no"
      ),
    model = NULL,
    predictor_levels = map(valid_predictors, ~levels(model_data[[.x]]))
  )
  
  # Only fit model if we have valid predictors
  if (length(valid_predictors) > 0) {
    formula <- reformulate(valid_predictors, "as.numeric(mh_trajectory)")
    results$model <- tryCatch(
      {
        glm(formula, data = model_data, family = poisson)
      },
      error = function(e) {
        message("Model fitting failed: ", e$message)
        return(NULL)
      }
    )
  }
  
  return(results)
}
# Suicide Risk Analysis
suicide_risk_analysis <- function(data) {
  data <- data %>%
    mutate(
      suicide_risk = case_when(
        suicide_ideation_12m == "Yes" ~ "High (12-month)",
        suicide_ideation_life == "Yes" ~ "Moderate (Lifetime)",
        TRUE ~ "None"
      ),
      suicide_risk = factor(suicide_risk, levels = c("None", "Moderate (Lifetime)", "High (12-month)"))
    )
  
  list(
    prevalence = data %>% 
      filter(immigration_status == "Immigrant") %>%
      tbl_summary(include = suicide_risk),
    model = MASS::polr(
      suicide_risk ~ professional_service_any + informal_support_any +
        family_support + barrier_stigma + years_in_canada,
      data = filter(data, immigration_status == "Immigrant"),
      Hess = TRUE)
  )
}

```

# Family Support Analysis
```{r results='asis'}
family_support_analysis <- function(data) {
  data <- data %>%
    mutate(
      family_support_strength = case_when(
        sr1_062 == "Very helpful" ~ "Strong",
        sr1_062 %in% c("Somewhat helpful", "Neither helpful nor unhelpful") ~ "Moderate",
        sr1_062 %in% c("Somewhat unhelpful", "Very unhelpful") ~ "Negative",
        sr1_60aa == "Yes" ~ "Present (unknown helpfulness)",
        TRUE ~ "None"
      )
    )
  
  list(
    desc = data %>%
      tbl_summary(include = c(family_support_strength, suicide_ideation_12m)),
    model = glm(
      suicide_ideation_12m ~ professional_service_any * family_support_strength +
        gender + age_group,
      data = data, family = binomial)
  )
}
```

#Descriptive Analysis

```{r results='asis'}
# Demographic characteristics
tbl_demographics <- data_processed %>%
  select(pop_centre_rural, marital_status, gender, age_group,
         immigration_status, years_in_canada, education) %>%
  tbl_summary(
    missing = "no",
    label = list(
      pop_centre_rural ~ "Population Centre",
      marital_status ~ "Marital Status",
      gender ~ "Gender"
    )
  ) %>%
  modify_caption("**Table 1: Demographic Characteristics**")

tbl_demographics
```

# Service Utilization Patterns

## Professional vs. Informal
```{r results='asis'}
tbl_service_compare <- data_processed %>%
  select(professional_service_any, informal_support_any) %>%
  tbl_summary(
    label = list(
      professional_service_any ~ "Professional Services",
      informal_support_any ~ "Informal Support"
    ),
    missing = "no"
  ) %>%
  modify_caption("**Table 2: Professional vs. Informal Service Utilization**")

# Detailed professional services
tbl_prof_detail <- data_processed %>%
  select(sr1_004a, sr1_004b, sr1_004c, sr1_004d, sr1_004e) %>%
  tbl_summary(
    label = list(
      sr1_004a ~ "Psychiatrist",
      sr1_004b ~ "Family Doctor",
      sr1_004c ~ "Psychologist",
      sr1_004d ~ "Nurse",
      sr1_004e ~ "Social Worker"
    ),
    missing = "no"
  ) %>%
  modify_caption("**Table 3: Detailed Professional Service Utilization**")

tbl_service_compare
tbl_prof_detail
```


# NEW: Enhanced Service Utilization by Demographics
```{r results='asis'}
demo_results <- demo_service_analysis(data_processed)
demo_results$gender
demo_results$age
demo_results$years_canada
demo_results$education

```

# Barriers Analysis
# Comprehensive barriers
```{r results='asis'}
data_processed <- data_processed %>%
  mutate(
    barrier_cost = factor(case_when(
      sr1_014e == "Yes" | sr1_024e == "Yes" | sr1_034e == "Yes" | sr1_044e == "Yes" | sr1_054e == "Yes" ~ "Yes",
      TRUE ~ "No"
    ), levels = c("No", "Yes")),
    barrier_stigma = factor(case_when(
      sr1_014j == "Yes" | sr1_024f == "Yes" | sr1_034f == "Yes" | sr1_044f == "Yes" | sr1_054f == "Yes" ~ "Yes",
      TRUE ~ "No"
    ), levels = c("No", "Yes")),
    barrier_access = factor(case_when(
      sr1_014h == "Yes" | sr1_024h == "Yes" | sr1_034h == "Yes" | sr1_044h == "Yes" | sr1_054h == "Yes" ~ "Yes",
      TRUE ~ "No"
    ), levels = c("No", "Yes"))
  )

tbl_barriers <- data_processed %>%
  select(starts_with("barrier_")) %>%
  tbl_summary(
    label = list(
      barrier_cost ~ "Financial Barriers",
      barrier_stigma ~ "Stigma Barriers",
      barrier_access ~ "Accessibility Barriers"
    ),
    missing = "no"
  ) %>%
  modify_caption("**Table 4: Prevalence of Service Access Barriers**")

tbl_barriers_immigration <- data_processed %>%
  select(starts_with("barrier_"), immigration_status, years_in_canada) %>%
  tbl_summary(by = immigration_status, missing = "no") %>%
  add_p() %>%
  modify_caption("**Table 5: Barriers by Immigration Status**")

tbl_barriers
tbl_barriers_immigration

# NEW: Professional-specific Barrier Analysis
barrier_results <- barrier_analysis(data_processed)
barrier_results$Psychiatrist
barrier_results$`Family Doctor`
barrier_results$Psychologist

```

#Suicide Ideation Analysis

# Prevalence
```{r results='asis'}
data_processed <- data_processed %>%
  mutate(
    suicide_ideation_life = labelled::to_factor(suicide_ideation_life),
    suicide_ideation_12m = labelled::to_factor(suicide_ideation_12m)
  )

tbl_suicide <- data_processed %>%
  select(suicide_ideation_life, suicide_ideation_12m) %>%
  tbl_summary(
    label = list(
      suicide_ideation_life ~ "Lifetime Suicidal Ideation",
      suicide_ideation_12m ~ "Past 12-Month Suicidal Ideation"
    ),
    missing = "no"
  ) %>%
  modify_caption("**Table 6: Prevalence of Suicidal Ideation**")

tbl_suicide_service <- data_processed %>%
  select(suicide_ideation_life, professional_service_any, informal_support_any) %>%
  tbl_summary(by = suicide_ideation_life, missing = "no") %>%
  add_p() %>%
  modify_caption("**Table 7: Service Use by Suicidal Ideation Status**")

tbl_suicide
tbl_suicide_service

```

# NEW: Longitudinal Mental Health Analysis

```{r results='asis'}
mh_results <- mh_trajectory_analysis(data_processed)
mh_results$desc_table
tbl_regression(mh_results$model, exponentiate = TRUE)
```



# NEW: Suicide Risk Stratification
```{r results='asis'}
risk_results <- suicide_risk_analysis(data_processed)
risk_results$prevalence
tbl_regression(risk_results$model)
```

#Family Support Analysis
```{r results='asis'}

data_processed <- data_processed %>%
  mutate(
    family_support = case_when(
      sr1_60aa == "Yes" | sr1_60ab == "Yes" | sr1_60ac == "Yes" | sr1_60ad == "Yes" ~ "Yes",
      TRUE ~ "No"
    ),
    family_support = factor(family_support, levels = c("No", "Yes"))
  )

tbl_family_support <- data_processed %>%
  select(family_support, suicide_ideation_life) %>%
  tbl_summary(by = suicide_ideation_life, missing = "no") %>%
  add_p() %>%
  modify_caption("**Table 8: Family Support by Suicidal Ideation Status**")

tbl_family_help <- data_processed %>%
  filter(family_support == "Yes") %>%
  select(sr1_062) %>%
  tbl_summary(
    label = list(sr1_062 ~ "Perceived Helpfulness of Family Support"),
    missing = "no"
  ) %>%
  modify_caption("**Table 9: Perceived Helpfulness of Family Support**")

tbl_family_support
tbl_family_help
```


# NEW: Comprehensive Family Support Analysis

```{r results='asis'}
family_results <- family_support_analysis(data_processed)
family_results$desc
tbl_regression(family_results$model, exponentiate = TRUE)
```

#Advanced Modeling

# Immigrant-specific model
```{r results='asis'}
immigrant_data <- data_processed %>%
  filter(immigration_status == "Immigrant") %>%
  select(
    suicide_ideation_life,
    gender,
    age_group,
    years_in_canada,
    professional_service_any,
    informal_support_any,
    barrier_cost,
    barrier_stigma,
    family_support
  ) %>%
  mutate(across(where(is.factor), droplevels)) %>%
  drop_na()

immigrant_model <- glm(
  suicide_ideation_life ~ gender + age_group + years_in_canada +
    professional_service_any + informal_support_any +
    barrier_cost + barrier_stigma + family_support,
  data = immigrant_data,
  family = binomial
)

model_results <- cbind(
  OR = exp(coef(immigrant_model)),
  exp(confint(immigrant_model))
) %>% 
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  mutate(across(where(is.numeric), ~round(.x, 2)))

model_results %>%
  gt() %>%
  fmt_number(columns = 2:4, decimals = 2) %>%
  cols_label(
    OR = "Odds Ratio",
    `2.5 %` = "Lower CI", 
    `97.5 %` = "Upper CI"
  ) %>%
  tab_header(
    title = "Predictors of Suicidal Ideation in Immigrants",
    subtitle = "Adjusted Logistic Regression Results"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = OR,
      rows = `2.5 %` > 1 | `97.5 %` < 1
    )
  )

```


```{r results='asis'}

# Model diagnostics
if(require(performance)) {
  check_model(immigrant_model)
  model_performance(immigrant_model) %>% print()
}
```

#Visualizations

```{r results='asis'}
# Service modality by age and immigration
ggplot(data_processed, aes(x = age_group, fill = any_video_service_use)) +
  geom_bar(position = "fill") +
  facet_wrap(~immigration_status) +
  labs(title = "Video Service Use by Age and Immigration Status",
       x = "Age Group", y = "Proportion Using Video Services") +
  scale_fill_brewer(palette = "Set1", name = "Used Video\nServices") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Mental health trajectory by family support
ggplot(data_processed %>% filter(!is.na(perceived_mh_compared_prev_year)), 
       aes(x = perceived_mh_compared_prev_year, fill = family_support)) +
  geom_bar(position = "fill") +
  labs(title = "Mental Health Trajectory by Family Support",
       x = "Perceived Change in Mental Health", y = "Proportion") +
  scale_fill_brewer(palette = "Set2", name = "Family Support") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

